Collocated Deduplication with Fault Isolation for Virtual Machine Snapshot Backup

### 2013-06-23 snapshot_system/failover.tex
revised analysis

We  compare the impact of losing $d$ machines 
to the the VC and VO approaches.  
The replication degree of the backup storage 
is $r$ for regular content blocks and $r=3$ is a typical setting in the distributed
file system~\cite{Hadoop,GFS}.
in The VC approach, a special replica degree $r_c$ used for CDS blocks where $r_c>r$. 
Noted that the storage cost for VO with full deduplication is $b_u *r$
while it is
\[ r*b*\delta(1 -   (\frac{k}{b_u})^{1-\alpha}) +k*(r-r_c).
\]

For each rank $i$ block which appears $C_i$ times  among $V$ virtual machines.
Assigning this block to virtual machines can be viewed as a classical ball-bin assignment
approximated by a Poisson distribution. Namely the virtual machines that share this block
is approximated as $1 -  e^{C_i/V}$.
Thus the average number of VMs that share a block is:
\[
(1-\delta) + \delta \sum_1^{u_b} \frac{1 -  e^{C_i/V}}{u_b}.
\]
Any failure of a block would impact the above number of virtual machines.
In VO, the number of VMs that share a block is:
\[
\sum_1^{u_b'} \frac{1 -  e^{C_i'/V}}{u_b'}.
\]

Next we discuss how a failure of $d$ machines impacts the VM snapshots in VC.
When $d<r$, there is no loss of data in the snapshot storage.
When $r_c> d \ge r$, some of data blocks in the storage are lost, and we compute the number of VMs that could
suffer the loss of their snapshots.
When $d \ge r_c$, some of CDS blocks are affected.
In VC, the blocks in CDS are stored in a container manner (superblock), and they are shared by many VMs.
The number of data containers stored with replication degree $r$ for
referenced by each VM is:
\[
N_1= \frac{b}{v*s} (1-\delta) +b/(v*s)\delta (1 -  (\frac{k}{b_u})^{1-\alpha})
\]
The number of CDS data containers stored with replication degree $r_c$ for
referenced by each VM is:
\[
N_2= \frac{b}{v*s}\delta (\frac{k}{b_u})^{1-\alpha}.
\]

When there are $r \le d<r_c$ machines failed and the probability of
a container failure depends on if all replicas of this container reside in the failed machines.
\begin{multline}
Probability (\mbox{ a VM fails in VC})\\
= 1- Probability (\mbox{ a VM has no data container loss}) \\
= 1-  Probability(\mbox{ A container has no data loss})^{N_1}
%\end{split}
\end{multline}

%Probablity(\mbox { no loss for this  block}) = 1 - Probability (\mbox{ this block in d failed machiens})
\begin{multline}
%\begin{split}
Probability(\mbox{A container no data loss})\\
= 1- Probability (\mbox{ this container is in d failed machines})\\
=  1- \frac{\binom{d}{r}} { \binom{p}{r} }.
%\end{split}
\end{multline}

When $r_c \leq d$, the CDS data containers in VC can also have a loss. 
\begin{multline}
Probability (\mbox{ a VM fails in VC})\\
= 1-  Probability(\mbox{ A container has no data loss})^{N_1}* \\
Probability(\mbox{ A CDS container has no data loss})^{N_2}\\
= 1 - 
(\frac{ \binom{d}{r}} { \binom{p}{r} })^{N_1} 
*
(\frac{ \binom{d}{r_c}} { \binom{p}{r_c} })^{N_2} 
%\end{split}
\end{multline}

By comparison, the probability that a VM has data loss is close to
1 once  there are  $d \ge r$ failed machines. That is because there are more data blocks
shared among VMs. Any failure of these blocks causes many more VMs to fail.

@@

To consider the reliability of VM backups in our model, we can consider the likely hood that a data chunk will fail,
given some number of storage node failures. Every time a filesystem chunk fails,
we say that we have lost data for that virtual machine, so it is no longer
available. (in reality it may be only one snapshot that is affected, but it is the user
who must decide which snapshots are important, so we consider the worst case). We use
filesystem chunks rather than data blocks as our unit of failure because the DFS keeps
filesystem chunks as its base unit of storage (in our case there are 16384 blocks in a
filesytem chunk, with 4KB block size and a 64MB chunk size).

We take the likelyhood of VM failure to be the same as the likelyhood of chunk failure,
so
\[
    P_{VMfailure}=\frac{\binom{d}{r}}{\binom{p}{r}}
\]

@@

in our current testbed we have measured the average number of links
per block (Figure~\ref{vo-links}) and number of links per chunk (Figure~\ref{vo-container-links})
in our dataset. For the complete dataset, we also found the average number of links to CDS chunks, which
was 11 using a 5\% CDS. From Figure~\ref{vo-container-links} you can see that as you add VMs, the average
number of links to a chunk increases. We expect that this trend will level out with larger datasets, but
should continue to increase at a slower rate. For the VC model, we have to consider both CDS chunks and 
Non-CDS chunks. In the VO model all data is shared, but in VC only the CDS data is shared, but at a much
higher rate. This gives a strong advantage to VC, in that for a large percent of the data (95-99\%), there
is only one VM pointing to each Non-CDS chunk. You can see the increase in reliability in Figure~\ref{fig-vm-availability}.
The figure shows that even as the number of links to a CDS chunk increase by 100 times, the availability only decreases
slightly, while in VO even small increases in $L$ causes a large decrease in availability. Figure~\ref{fig-cds-replication}
shows the advantages of increasing the replication factor for CDS blocks. These figures together show
the advantages of the VM-centric model, and the advantages that separating the replication factor for popular blocks
can have on reliability.

@@

We can build a bipartite graph representing the association from unique data blocks
to their corresponding VMs. An association edge is  drawn  from a data block  to a VM 
if this block is used by this VM, 
The average incoming degree of VM nodes is equal to
the average outgoing degree of data blocks, which is $L_c$.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.45\textwidth]{shared.pdf}
\caption{Assoication of data blocks to VMs.}
\label{fig:shared}
\end{figure}

>>>>

We  compare the impact of losing $d$ machines 
to the the VC and VO approaches.  
The replication degree of the backup storage 
is $r$ for regular content blocks and $r=3$ is a typical setting in the distributed
file system~\cite{Hadoop,GFS}.
In the VC approach, a special replica degree $r_c$ used for CDS blocks where $r_c>r$. 
Notice that the ratio of non CSD data size vs CDS data size for each VM is
\[
\sigma=  \frac{b*\delta(1 -   (\frac{k}{b_u})^{1-\alpha})} {k}.
\]

Thus  storage cost for VO with full deduplication is $b_u *r$ and for VC, it is
\[ k*r_c  + k*\sigma *r
\]

In our experiment with Alibaba data, the ratio $\sigma$ 
is 162. Thus allocation of extra replicas for CDS only introduces a small amount of extra space
cost.  Figure~\ref{fig:storageCDS} shows the storage cost ratio of VC and VO when 
$r$=3, and $r_c$ varies from 3 to 10.
The result shows that the storage cost for adding extra replication for CDS
is insignificant.
Now we compare the fault resilience of VC and VO.

\comments{
For each rank $i$ block which appears $C_i$ times  among $V$ virtual machines.
Assigning this block to virtual machines can be viewed as a classical ball-bin assignment
approximated by a Poisson distribution. Namely the virtual machines that share this block
is approximated as $1 -  e^{C_i/V}$.
Thus the average number of VMs that share a block is:
\[
(1-\delta) + \delta \sum_1^{u_b} \frac{1 -  e^{C_i/V}}{u_b}.
\]
Any failure of a block would impact the above number of virtual machines.
In VO, the number of VMs that share a block is:
\[
\sum_1^{u_b'} \frac{1 -  e^{C_i'/V}}{u_b'}.
\]

Next we discuss how a failure of $d$ machines impacts the VM snapshots in VC.
When $d<r$, there is no loss of data in the snapshot storage.
When $r_c> d \ge r$, some of data blocks in the storage are lost, and we compute the number of VMs that could
suffer the loss of their snapshots.
When $d \ge r_c$, some of CDS blocks are affected.
In VC, the blocks in CDS are stored in a container manner (superblock), and they are shared by many VMs.
The number of data containers stored with replication degree $r$ for
referenced by each VM is:
\[
N_1= \frac{b}{v*s} (1-\delta) +b/(v*s)\delta (1 -  (\frac{k}{b_u})^{1-\alpha})
\]
The number of CDS data containers stored with replication degree $r_c$ for
referenced by each VM is:
\[
N_2= \frac{b}{v*s}\delta (\frac{k}{b_u})^{1-\alpha}.
\]
}

%Thus the number of VMs using a block =    $(1-s_c)  + s_c *L$.
In characterizing the reliability of VM backups in our model, 
we consider the likely hood that a file system block fails,
given some number of storage node failures. 
Every time a filesystem block fails,
we say that we have lost data for that virtual machine, so it is no longer
available. 
(in reality it may be only one snapshot that is affected, but it is the user
who must decide which snapshots are important, so we consider the worst case). 
We use filesystem blocks rather than a deduplication
data chunk as our unit of failure because the DFS keeps
filesystem blocks as its base unit of storage (in our case there are 
16384 blocks in a filesytem block on average, with 4KB block size and 
a 64MB block size).

To  compute the probablity of losing a virtual machine, 
we estimate the number of file system blocks per VM in each apporach.
We can build a bipartite graph representing the association from unique file system blocks
to their corresponding VMs. An association edge is  drawn  from a file block  to a VM 
if this file is used by this VM. 
For VC, let $N_1$ be the averge number of file system blocks for non-CDS data for each VM
and  $N_2$ be the averge number of file system blocks for CDS data for each VM.
For VO, let $N_o$ be the averge number of file system blocks for each VM
and  let $V_o$ be the averge number of VMs shared by each file system block.
Figure~\ref{fig:shared} illustates the bipartite association.
\begin{figure}[htbp]
\centering
\includegraphics[width=0.45\textwidth]{shared.pdf}
\caption{Assoication of data blocks to VMs.}
\label{fig:shared}
\end{figure}

Each non-CDS file system block is associated with one VM while CDS file system blocks are
shared among VMs. 
\[
V *N_1 *s  = pD  \frac{\delta} {\delta +1}
\]
For CDS file blocks, there are at most $V$ VMs sharing, thus
\[
V *N_2 *s  \leq pD  \frac{1} {\delta +1} *V
\]
For the VO approach, 
\[
V *N_o *s  = pD  V_o.
\]
Then
\[
N_1= \frac{pD \delta} {V s (\delta +1)},\; 
N_2 \leq \frac{pD } {s (\delta +1)}, \; 
\mbox{and} N_o = \frac{pD V_o } {s V}.
\]
Since each file block (with default size $s=64MB$) contains many chunks (on average 4KB),
each file block contains the hot low-level chunks shared by many VMs, and it also contains
rare chunks which are not shared.
Figure~\ref{fig:VOshared} shows the number of VMs shared by each file block.
In our experiment, we find that $V_o \approx 0.2 V$ when backuping up VMs one by one.
We can observe that $N_1 +N_2 << N_o$.
If  the backup for multiple VMs is conducted concurrently, there would be more
VMs sharing for each file block on average.

The availability of is a VM is the likelyhood that
there is no  data loss for all its file blocks.
With replication degree $r$, the likelyhood of
a file block block is the probability that  
all of its replicas appear in $d$ failed machines. 
Namely, $\binom{d}{r}/ \binom{p}{r}$.

When there are $r \le d<r_c$ machines failed, 
%a file block failure depends on if all replicas of this container reside in the failed machines.
the avaiability of a VM in the VC approach is
\[
(1-\frac{\binom{d}{r}} { \binom{p}{r} })^{N_1}.
\]
%\begin{multline}
%= Probability (\mbox{ There is no file block loss in this VM}) \\
%= Probability(\mbox{ A file block has file data loss})^{N_1}\\
%= (1- Probability (\mbox{all replicas of its block fail}))^{N_1}\\
%\end{multline}
%\end{split}

When $r_c \leq d$, the CDS file block in VC can also have a loss.
The availability of  a VM in the VC approach is
\[
(1-\frac{ \binom{d}{r}} { \binom{p}{r} })^{N_1} 
*
(1-\frac{ \binom{d}{r_c}} { \binom{p}{r_c} })^{N_2}.
\]
%= Probability(\mbox{ No non-CDS file block loss})^{N_1}
%Probability(\mbox{ No CDS file block loss})^{N_2}\\
%\begin{multline}
%\end{split}
%\end{multline}

The availability of its VM  mbox in the VO approach is
\[ 
(1-\frac{ \binom{d}{r}} { \binom{p}{r} })^{N_o}. 
\]

%By comparison, the probability that a VM has data loss is close to
%1 once  there are  $d \ge r$ failed machines. That is because there are more data blocks
%shared among VMs. Any failure of these blocks causes many more VMs to fail.

@@

\comments{
We take the likelyhood of VM failure to be the same as the likelyhood of chunk failure,
so
\[
    P_{VMfailure}=\frac{\binom{d}{r}}{\binom{p}{r}}
\]

@@

in our current testbed we have measured the average number of links
per block (Figure~\ref{vo-links}) and number of links per chunk (Figure~\ref{vo-container-links})
in our dataset. For the complete dataset, we also found the average number of links to CDS chunks, which
was 11 using a 5\% CDS. From Figure~\ref{vo-container-links} you can see that as you add VMs, the average
number of links to a chunk increases. We expect that this trend will level out with larger datasets, but
should continue to increase at a slower rate. 
For the VC model, we have to consider both CDS chunks and 
Non-CDS chunks. In the VO model all data is shared, but in VC only the CDS data is shared, but at a much
higher rate. This gives a strong advantage to VC, in that for a large percent of the data (95-99\%), there
is only one VM pointing to each Non-CDS chunk. You can see the increase in reliability in Figure~\ref{fig-vm-availability}.
}

The figure shows that even as the number of VMs sharing a CDS file block 
increase by 100 times, the availability only decreases
slightly.
The key factor placed is that  $N_1 +N_2 << N_o$.
The VM-centric approach packs  data blocks under one VM as much as possible
and performs deduplication mainly within the same VM. The sharing part in CDS is replicated
with a higher degree, which also signficantly increses the availability even a CDS file blcok
is shared by every VM.

% while in VO even small increases in $L$ causes a large decrease in availability. 
Figure~\ref{fig-cds-replication}
shows the advantages of increasing the replication factor for CDS blocks. These figures together show
the advantages of the VM-centric model, and the advantages that separating the replication factor for popular blocks
can have on reliability.

@@

### 2013-06-23 snapshot_system/snapshot.tex
revised analysis

\begin{table}[ht]
\centering
\begin{tabular}{|p{1.25cm}|p{6.5cm}|}
\hline
$b$ &  the total amount of data blocks\\ 
\hline
$b_u$ &  the total amount of unique fingerprints after inner VM  deduplication\\
\hline
$C_i$ &  the frequency for the $i$th most popular fingerprint\\
\hline
$\delta$ &  the percentage of duplicates detected in inner VM deduplication\\
\hline
$p$ & the number of nodes in the cluster\\
\hline
$D$ & the amount of unique data on each node\\
\hline
$B$ & the average block size. Our setting is  4K.\\
\hline
$s$ & the average size of data containers stored in the distributed file system . Our setting is  64MB.\\
\hline
$m$ & memory size on each node used by VC\\ 
\hline
$F$ & the size of an polular data index entry\\
\hline
\end{tabular}
\caption{Modeling  symbols.}
\label{tab:symbol}
\end{table}

@@

The  $k$ most popular fingerprints can cover the following number of blocks.
\[
C_1 \sum_{i=1}^{k}\frac{1}{i^\alpha} \approx  
C_1 \int_{1}^{k}\frac{1}{x^\alpha} dx  \approx  C_1\frac{  k^{1-\alpha}} {1-\alpha}.
\]

Deduplication efficiency of VC using top $k$ popular blocks
is the percentage of duplicates that can be detected:  
\begin{equation}
\begin{split}
e_k &= 
\frac{ b(1-\delta) + C_1\frac{  k^{1-\alpha}} {1-\alpha}} 
{b(1-\delta)  + sb - b_u }\\
&= 
\frac{ (1-\delta) + s (\frac{k}{b_u})^{1-\alpha}}
{ 1- \frac{b_u}{b} }.
\end{split}
\end{equation}

@@

The overall deduplication effiency is
\[
\frac{ (1-\delta) + \delta (\frac{m*B}{D*F})^{1-\alpha}}
{ 1- \frac{b_u}{b} }.
\]
where $(\frac{m*B}{D*F})^{1-\alpha}$ represents the percentage of the remaining blocks detected as duplicates
after inner VM deduplication. 
%Then the deduplication efficiency of CDS approach using top k most popular fingerprints is:
%\begin{equation}
%  e_k = (\frac{m*B}{F*D})^{1-\alpha}
%\end{equation}

@@

\begin{figure}[ht]
  \centering
    \begin{tikzpicture}
            \begin{axis}[
            %title={CDS Coverage},
            xlabel={Total Data Stored (GB)},
            ylabel={CDS Coverage (\%)},
            %extra y ticks={4.5,5.5,6.5} %to add extra ticks
            %legend pos=outer north east,
            legend columns=2,
            legend style={
                at={(0.5,-0.2)},
            anchor=north}
            ]
            \addplot[color=blue,mark=*] table[x=InputSize,y=A1] {figures/cds_coverage.txt};
            \addplot[color=blue,mark=square*] table[x=InputSize,y=A2] {figures/cds_coverage.txt};
            \addplot[color=blue,mark=triangle*] table[x=InputSize,y=A4] {figures/cds_coverage.txt};
            \addplot[densely dashed,color=red,mark=*] table[x=InputSize,y=T1] {figures/cds_coverage.txt};
            \addplot[densely dashed,color=red,mark=square*] table[x=InputSize,y=T2] {figures/cds_coverage.txt};
            \addplot[densely dashed,color=red,mark=triangle*] table[x=InputSize,y=T4] {figures/cds_coverage.txt};
            \legend{Measured (1\%CDS),Measured (2\%CDS),Measured (4\%CDS),Predicted (1\%CDS),Predicted (2\%CDS),Predicted (4\%CDS)};
            \end{axis}
    \end{tikzpicture}
  \caption{fixed-alpha predicted vs. actual CDS coverage as data size increases.}
  \label{fig-vo-links}
\end{figure}

>>>>

\begin{table}[ht]
\centering
\begin{tabular}{|p{1.25cm}|p{6.5cm}|}
\hline
$b$ &  the total amount of data blocks\\ 
\hline
$b_u$ &  the total amount of unique fingerprints after inner VM  deduplication\\
\hline
$C_i$ &  the frequency for the $i$th most popular fingerprint\\
\hline
$\delta$ &  the percentage of duplicates detected in inner VM deduplication\\
\hline
$\sigma$ &  the percentage of $b_u$ unique items  stored in CDS.\\
\hline
$p$ & the number of machines in the cluster\\
\hline
$D$ & the amount of unique data on each machine\\
\hline
$B$ & the average block size. Our setting is  4K.\\
\hline
$s$ & the average size of data containers stored in the distributed file system . Our setting is  64MB.\\
\hline
$m$ & memory size on each node used by VC\\ 
\hline
$F$ & the size of an polular data index entry\\
\hline
\end{tabular}
\caption{Modeling  symbols.}
\label{tab:symbol}
\end{table}

@@

The  $k$ most popular fingerprints can cover the following number of blocks after inner VM 
deduplication.
\[
C_1 \sum_{i=1}^{k}\frac{1}{i^\alpha} \approx  
C_1 \int_{1}^{k}\frac{1}{x^\alpha} dx  \approx  C_1\frac{  k^{1-\alpha}} {1-\alpha}.
\]

Deduplication efficiency of VC using top $k$ popular blocks
is the percentage of duplicates that can be detected:  
\begin{equation}
\begin{split}
e_k &= 
\frac{ b(1-\delta) + C_1\frac{  k^{1-\alpha}} {1-\alpha}} 
{b(1-\delta)  + \delta b - b_u }\\
&= 
\frac{ (1-\delta) + \delta  (\frac{k}{b_u})^{1-\alpha}}
{ 1- \frac{b_u}{b} }.
\end{split}
\end{equation}

@@

The overall deduplication effiency of VC is
\[
\frac{ (1-\delta) + \delta (\frac{m*B}{D*F})^{1-\alpha}}
{ 1- \frac{b_u}{b} }.
\]
where $(\frac{m*B}{D*F})^{1-\alpha}$ represents the percentage of the remaining blocks detected as duplicates
after inner VM deduplication. 
%Then the deduplication efficiency of CDS approach using top k most popular fingerprints is:
%\begin{equation}
%  e_k = (\frac{m*B}{F*D})^{1-\alpha}
%\end{equation}

@@

\begin{figure}[ht]
  \centering
    \begin{tikzpicture}
            \begin{axis}[
            %title={CDS Coverage},
            xlabel={Total Data Stored (GB)},
            ylabel={CDS Coverage (\%)},
            %extra y ticks={4.5,5.5,6.5} %to add extra ticks
            mark options=solid,
            %legend pos=outer north east,
            legend columns=2,
            legend style={
                at={(0.5,-0.2)},
            anchor=north}
            ]
            \addplot[color=blue,mark=*] table[x=InputSize,y=A1] {figures/cds_coverage.txt};
            \addplot[color=blue,mark=square*] table[x=InputSize,y=A2] {figures/cds_coverage.txt};
            \addplot[color=blue,mark=triangle*,mark size=3] table[x=InputSize,y=A4] {figures/cds_coverage.txt};
            \addplot[densely dashed,color=red,mark=*] table[x=InputSize,y=T1] {figures/cds_coverage.txt};
            \addplot[densely dashed,color=red,mark=square*] table[x=InputSize,y=T2] {figures/cds_coverage.txt};
            \addplot[densely dashed,color=red,mark=triangle*] table[x=InputSize,y=T4] {figures/cds_coverage.txt};
            \legend{Measured (1\%CDS),Measured (2\%CDS),Measured (4\%CDS),Predicted (1\%CDS),Predicted (2\%CDS),Predicted (4\%CDS)};
            \end{axis}
    \end{tikzpicture}
  \caption{fixed-alpha predicted vs. actual CDS coverage as data size increases.}
  \label{fig-vo-links}
\end{figure}
